Simulating an Interplanetary Internet

### 2024-04-10 paper/paper.tex
fix typo
fix some details

\title{Simulating an Interplanetary Internet}
\author{Cody Auch, Mark Lysack, Jacob Janzen}
\date{\today}

@@

\begin{abstract}
  This paper discusses interplanetary networks, the challenges that can arise
  from them that make them a unique challenge compared to shorter-range computer
  networks, and how our current network protocols can be applied in this
  application. In addition, we look at how one can use conventional network
  simulators to test interplanetary network implementations. We used the network
  simulator NS-3 ~\cite{ns-3} to test various transport layer protocols against
  each other in a network topology that we designed in the network simulator to
  emulate long-distance communications. Using this network simulation, we were
  able to compare UDP, TCP, and TCP New Reno protocols against each other to see
  how they each performed across such large distances with high delays. UDP
  preformed the best with New Reno and TCP closely paired together.
\end{abstract}

@@

The first assumption made was that all orbits are circular. Realistically,
orbits are oval shaped, with two focal points. A result of this assumption is
the elimination of any difference in orbital speed due to Kepler's third law.
Normally a planet will accelerate or decelerate depending on where it is in its
orbit around the sun. The orbital radius was chosen by taking the average
orbital radius of the real life celestial bodies. For example Earth's average
orbital radius is 149.6 billion meters~\cite{nasa_earth}, and therefore orbits
around the sun in a circular orbit of radius 149.6 billion meters in our model.
Another result of this assumption is that orbits are considered to be centered
around the body which it orbits. The Mars orbit has a focal point that is
further in space, making it more oval shaped than the earth
orbit.~\cite{nasa_mars} While this could make changes to a simulated network in
some very specific situations, the primary behaviour of a solar system is
maintained, like retrograde orbits, periodical alignment, predictable behaviour,
and recursive orbits of sub orbiting object.

@@

The graph shows a large amount of sensitivity to the motion of the planets which
can be seen by the curvature of the graph. At time around 28000 seconds, the
topology suddenly changed. Caused by a satellite blocking the path between Mars
and Earth. This caused a secondary, less efficient path to be taken. This shows
why there are several ``layers'' of time which it took. They likely represent the
different paths which the message took before being sent to its destination.
Each quanta in the time it took to complete the message represents the smaller
jump it takes to get from a satellite to its orbiting entity. The increase in
time likely represents that while the direct Earth and Mars link was down the
smaller satellites were actually orbiting away from the message destination. UDP
messages spent much more time being queued than the other messages. This may
imply that messages were being transferred back and forth between two closely
orbiting objects, which could explain the many different layers which it can
take for a message to be sent. It could also be that since the average message
is larger than TCP or New Reno it takes longer for the message to be sent and it
is transmitting.

@@

In this paper we outlined how we created a model of an interplanetary internet
(IPN) using a physics engine which fed into a NS-3 simulation. The physics
simulation used a simplified model of orbiting planets and satellites as well as
physics calculations to create data which would be the output for any input time
$t$. The NS-3 model used a point to point base which programmatically took the
data from the physics engine and created nodes which acted as routers. A
simulation was ran which transfered data between Earth and Mars, using
satellites and other entities to relay messages around blocked connections.
Analysis was done on three different network protocols used in the IPN, that is
UDP, TCP and New Reno. UDP was found to have the most throughput out of all of
the three protocols, while New Reno reduced the amount of time a message was
queued.

>>>>

\title{Simulating an Interplanetary Internet}
\author{Cody Auch, Jacob Janzen, Mark Lysack}
\date{\today}

@@

\begin{abstract}
  This paper discusses interplanetary networks, the challenges that can arise
  from them that make them a unique challenge compared to shorter-range computer
  networks, and how our current network protocols can be applied in this
  application. In addition, we look at how one can use conventional network
  simulators to test interplanetary network implementations. We used the network
  simulator NS-3~\cite{ns-3} to test various transport layer protocols against
  each other in a network topology that we designed in the network simulator to
  emulate long-distance communications. Using this network simulation, we were
  able to compare UDP, TCP, and TCP New Reno protocols against each other to see
  how they each performed across such large distances with high delays. UDP
  preformed the best with New Reno and TCP closely paired together.

  Our implementation can be found
  at \url{https://github.com/codyauch/IPN-Project}. The physics simulation was
  mostly designed by Mark and Cody. The network simulation in NS-3 was done by
  Jacob, and the analysis was done by Mark.
\end{abstract}

@@

The first assumption made was that all orbits are circular. Realistically,
orbits are oval shaped, with two focal points. A result of this assumption is
the elimination of any difference in orbital speed due to Kepler's second law.
Normally a planet will accelerate or decelerate depending on where it is in its
orbit around the sun. The orbital radius was chosen by taking the average
orbital radius of the real life celestial bodies. For example Earth's average
orbital radius is 149.6 billion meters~\cite{nasa_earth}, and therefore orbits
around the sun in a circular orbit of radius 149.6 billion meters in our model.
Another result of this assumption is that orbits are considered to be centered
around the body which it orbits. The Mars orbit has a focal point that is
further in space, making it more oval shaped than the earth
orbit.~\cite{nasa_mars} While this could make changes to a simulated network in
some very specific situations, the primary behaviour of a solar system is
maintained, like retrograde orbits, periodical alignment, predictable behaviour,
and recursive orbits of sub orbiting object.

@@

We believe that UDP sends a large number of messages at the start that are all queued, processed, and sent. After those messages are sent, the next batch is prepared immediately. This curve could be caused by the queuing delay from the large number of messages that are sent. UDP does not wait for messages to be received and will therefore create a large backlog which slows everything down when dequeuing messages.

@@

In this paper we outlined how we created a model of an interplanetary internet
(IPN) using a physics engine which fed into a NS-3 simulation. The physics
simulation used a simplified model of orbiting planets and satellites as well as
physics calculations to create data which would be the output for any input time
$t$. The NS-3 model used a point to point base which programmatically took the
data from the physics engine and created nodes which acted as routers. A
simulation was ran which transferred data between Earth and Mars, using
satellites and other entities to relay messages around blocked connections.
Analysis was done on three different network protocols used in the IPN, that is
UDP, TCP and New Reno. UDP was found to have the most throughput out of all of
the three protocols, while New Reno reduced the amount of time a message was
queued.
